%Analýza a návrh øešení konkrétního problému

%Analýza a návrh øešení problému z oblasti úzce související se 
%softwarovým èi datovým inženýrstvím nebo systémovým prostøedím.
%Výsledkem je výzkumná zpráva a prototypové øešení.



%Vypracování musí demonstrovat pochopení problému. Tedy práce by
%nemìla pouze slepì následovat první zjevnou cestu k øešení, èi
%cestu naznaèenou v zadání. Autor musí ze všeho nejdøíve ukázat, že
%chápe, jaký problém má øešit, a že je schopen vysvìtlit, èeho a
%proè chce dosáhnout.

%Text práce má dokumentovat postup vypracování øešení tak, aby bylo
%možné posoudit, zda si autor poèínal vhodnì, a zda si byl vìdom
%souvislostí, které jeho vypracování ovlivòují. Tedy je tøeba
%systematicky a pøehlednì vysvìtlit a zdùvodnit dùležitá rozhodnutí
%uèinìná pøi vypracování práce. Text práce má popisovat zvolené
%øešení strukturovanì shora dolù, a zabíhat do detailu pouze v
%pøípadech nestandardních postupù nebo rozhodnutí.

%Pokud práce z prostorových èi kapacitních dùvodù neobsahuje
%nejlepší možné øešení problému, mìla by vysvìtlit, jak by
%optimální øešení mohlo vypadat.

%Hrubá osnova výzkumné zprávy:

%1.         Úvod charakterizující kontext problému - introduction
%2.         Pøesné vymezení problému 
%3.         Pøesné vymezení cílù návrhu
%4.         Struktura práce ve vztahu k vytyèeným cílùm
%5.         Analýza problému z rùzných hledisek
%6.         Návrh øešení s diskusí alternativ
%7.         Rozpracování vybraného pøístupu
%8.         Popis prototypové implementace
%9.         Zhodnocení splnìní cílù a srovnání s jinými známými pøístupy
%10.        Závìr se struèným shrnutím výsledkù a pøínosu
%11.        Seznam citované literatury

%Implementace musí být odevzdána jednak v podobì zdrojových souborù a všech dat potøebných k jejímu pøekladu a spuštìní, jednak v již pøeložené a spustitelné podobì. Na zdrojových souborech musí být jasnì patrné, které èásti implementace jsou autorovým pùvodním dílem, a které jsou pøevzaté (generovaný kód, knihovny ...).

%Spustitelná podoba práce musí demonstrovat funkènost implementace. Pokud je k takové demonstraci potøeba netriviální prostøedí èi nastavení, autor práce jí musí na požádání pøedvést vedoucímu a oponentovi, pøípadnì také hodnotící komisi.

\documentclass [10pt]{report}

\title{The Tool for Modeling of Evolution of the Artificial Life}
\author{Iva Bartunkova}

\usepackage{url}

\begin{document}

\maketitle

\begin{abstract}
Software simulator allowing experiments with the evolution of simple organisms.
\end{abstract}

\chapter{Introduction}

%*************3 pages about 1.Artificial life - definition, history, contemporary course 2. Brief context of my thesis*******************************

%******************Definition of Artificial life*********************
Artificial Life, (commonly \textbf{Alife} or \textbf{alife}) is a field of study and art form that examines systems related to life, its processes and its evolution through simulations using computer models, robotics, and biochemistry (called "soft", "hard", and "wet" approaches respectively \cite{Bedau}).

Because of its predominance within the field, the term "Artificial Life" is often used to specifically refer to soft alife.%source??

The term Artificial life was first coined by Christopher Langton in the late 1980s at the first "International Conference on the Synthesis and Simulation of Living Systems" (otherwise known as Artificial Life I) at the Los Alamos National Laboratory in 1987. He envisioned a study of life as it could be in any possible setting.

Artificial life studies "natural" life by attempting to recreate biological phenomena from scratch within non-living media like computers or RNA structures of molecules. Alife complements the traditional analytic approach of traditional biology with a synthetic approach in which, rather than studying biological phenomena by taking apart living organisms to see how they work, one attempts to put together systems that behave like living organisms.

The process of synthesis has been an extremely important tool in many disciplines. Synthetic chemistry - the ability to put together new chemical compounds not found in nature - has not only contributed enormously to our theoretical understanding of chemical phenomena, but has also allowed us to fabricate new materials and chemicals that are of great practical use for industry and technology.

Artificial life amounts to the practice of "synthetic biology" and, by analogy with synthetic chemistry, the attempt to recreate biological phenomena in alternative media will result in not only better theoretical understanding of the phenomena under study, but also in practical applications of biological principles in the technology of computer hardware and software.\cite{Langton} 

The seminal novelty of ALife lies in its synthetic approach. Whereas traditional research is essentially analytic, breaking down complex systems into basic components, Alife attempts to construct complex systems from elemental units.\cite{BornhoffenLataud}

For example, answering of questions like how the simple rules of Darwinian evolution lead to high-level structure, or the way in which the simple interactions between ants and their environment lead to complex trail-following behavior promises to provide novel solutions to complex real-world problems, such as disease prevention, stock-market prediction, and data-mining on the Internet.\cite{AdamiBrown}

Artificial life can be split into various subfields according to chosen substrat of the system (digital, mechanical or test-tube) as well as according to aspect of natural life, that is in the focus of the research.


%replace this to another chapter!!
%*************Evolutionary computation: if we choose evolution of biological life as the topic of our artificial life exploration and simulate it using computers.
One of terms related to Artificial life is Evolutionary computation. It is a general term for several computational techniques which are based to some degree on the evolution of biological life in the natural world and simulated using computers. Evolutionary computation thus uses techniques of natural world and applies them using computers in other fields of interest. The most widely used form of evolutionary computation are Genetic Algorithms. Evolution techniques are often focused on evolution of motion, shape, visual art and music. 
 
%************History of Alife as a science
%The only source: Wikipedia - History of artificial life. However, citations are poor in this article. 
%Exists:Artificial Life and the Sciences of Complexity: History and Future, 	Jean-Claude Heudin. But it is not accessible for free.
%1hour of searching. Lets stop it.

%************Contemporary course of this science


%***********This thesis is about -> software
This thesis deals with digital substrate, so it is bounded within software approach to Artificial life. Because of the fact that Artificial life uses synthetic methods, the logical realization of software based Artificial life are software simulators that perform the construction of complex systems from elemental units.

%***********Target of this work: 1.set of features 2.simulator 3.experiments
Target of this thesis is to introduce a simulator named Abeetles, that mimics life of simple organisms and is concerned with evolution of deliberately chosen features and behavior. The system avails experiments with evolution of the organisms under different conditions. Furthemore, several experiments performed with the system are described and a few other proposed.


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Artificial life simulators and position of Abeetles among them} 
%Pøesné vymezení problému
%5 pages about precise definition of the problem - description of chosen existing brother programs and comparison and choice of my objectives. I will use the Porovnani.xls 
%******************Context of my thesis - cathegories of Alife simulators and ranking of my work. Short about existing brother programs and programs that inspired me.
Software simulators of Artificial life can be ranked in many ways. This thesis will expand on three classification possibilities. First, cathegorisation according to the method of creature definition will be used. Second, simulators will be distinguished with respect to the feature or features of Artificial life, that they attempt to simulate. Third, the criterion of purpose of the simulator and target group of users will be used. Four section places Abeetles into thse classification systems and thereby specifies sphere of Artificial life that will be dealt in this thesis

\section{Classification by creature definition}
Individual agents are in existing systems modeled and constructed in many different ways. Most of them can be classified in the following way:
\begin{itemize}
    \item \textbf{Program Based} - In program based simulators an individual is represented by a program, which substitutes biological DNA. Language of the program is usually Turing complete. Assembly derivatives are the most common languages used. Tom Ray's Tierra is the most famous example of a program based simulator.
    \item \textbf{Module Based} - An agent in a module based system is a composition of individual modules. These modules modify the creature's behaviors and characteristics either directly, by hard coding into the simulation (leg type A increases speed and metabolism), or indirectly, through the emergent interactions between a creature's modules (leg type A moves up and down with a frequency of X, which interacts with other legs to create motion). Generally these are simulators which emphasize user creation and accessibility over mutation and evolution.  
    \item \textbf{Parameter Based} - If an organism is generally constructed with defined and fixed behavior, that is controlled by various parameters that mutate, the system is referred as a parameter based. hat is, each organism contains a collection of numbers or other finite parameters. Each parameter controls one or several aspects of an organism in a well defined way.
    \item \textbf{Neural Network Based} - These simulations have creatures that learn and grow using neural networks or a close derivative. Emphasis is often, although not always, more on learning than on natural selection.
\end{itemize}

\section{Classification by simulated phenomenon}
Another grouping of simulators can be done according to phenomenon, that they try to simulate. \url{http://vlab.infotech.monash.edu.au/}
\begin{itemize}
  \item \textbf{Networks} - Complex natural systems are simulated according to network theory (or diktyology) as networks or graphs and their statistical and topological properties are analyzed. 
  \item \textbf{Nonlineality} - Nonlineal systems are systems that cannot be mathematically described as the sum of their components. While certain assumptions can be made for linear systems, that often make the mathematical modelling of such systems easy, mathematical modelling of nonlineal systems is often very difficult or impossible. As a result, nonlineal systems are often studies through use of simulations.
  \item \textbf{Swarms} - A swarm is a group of independent agents that gather together in order to collectively carry out a certain task. Typically, each agent exhibits a very simple behaviour pattern that is influenced by direct or indirect interactions with other swarm members. As a result, the swarm as a whole may exhibit complex and intelligent behaviour patterns. In nature, swarms can be observed in social insects, fish schools, but also in primitive single cell organisms.
  \item \textbf{Evolution} - Evolution is the process of development or grows by accumulation of small advantageous changes. The study of all forms of evolutionary processes is one of the primary goals of ALife. This includes the study of biological evolution of species as well as other evolutionary processes in natural, artificial and social systems.
  \item \textbf{Cellular Automata} - A cellular automaton (CA) is a discrete model that consists of a cell-grid, where each cell is in one of a number of well-defined states. The grid can be in any finite number of dimensions, though the mostly studied CAs have one or two dimensions. The time in the model is discrete and the state of each cell C at time t depends only on the cell-states within some well-defined neighbourhood of C at time t-1. The update function (also called the update rule) is a table that defines the state for each cell at time t as a function of the state of the neighbourhood of C at time t-1. Cellular automaton is not really phenomenon of the natural life, it is a  discrete model of mutual influence of neighboring elements, which can be in the nature observed. 
\end{itemize}  

\section{Classification by purpose}
Apparently, the main purpose of Artificial life simulators is to simulate artificial life. The target can be the complexity of simulation as well as the concentration on a small selection of its attributes. As the simulator is a software system, it is designed for a certain group of users and expected usage. And decision in this field influence interface, adjustability and output of the program. Also availability of the system is related to the purpose. 
\begin{itemize}
  \item \textbf{Games} Frequent purpose of simulators is entertainment, because game industry is constantly researching for new ideas to animate artificial characters. Simulators from this class are usually supplied with attractive user interface, but neither artificial life techniques and settings nor source corde are accessible to examination. The game Creatures is a well-known example of an artificial life computer program series, created in the mid-1990s by English computer scientist Steve Grand. The program is regarded as an important breakthrough in the advancement of artificial life research.
  \item \textbf{Scientific simulators} A simulator can be designed primarily for use as a platform in Artificial Life research. Such programs allow to perform experiments in certain subfields of Alife, e.g. evolutionary dynamics, theoretical biology etc. They are usually highly adjustable and configurable and afford opportunities to gather high-quality statistics. The grafical output is not expected to be the most important feature, in comparison with games. They are often open source. Tom Ray's system Tierra and Avida from Devolab are such simulations. 
  \item \textbf{Simulators for teaching purposes}
Efforts of creators of a simulator can be also concencentrated on an idea of demonstration of natural life processes using computer with the objective that through experimentation and interactive play users learn underlying patterns of life. Intriguing example is the simulator Mitozoos, where through mutation of spider-like agents can anybody explore relationship between genetic code and life. 
\item \textbf{Related systems}
Many systems are closely related to Artificial life simulators, but they are not considered as Artificial life systems, because they only use techniques, that originate from Alife like genetic algorithms or ant colony optimization. Their purpose can be various. The primary difference lies in the fact that these simulators explicitly define the fitness of an agent by its ability to solve a problem, instead of its ability to find food, reproduce, or avoid death. 
\end{itemize}

\section{Classification of Abeetles}
As far as definition of agents is concerned, target class of Abeetles are parameter based systems. Space of possibilities of Abeetles is bounded by finite number of parameters and Abeetles avails to explore what implies which setting of parameters in the world of this system.

In classification by simulated phenomenon Abeetles can be ranked among systems concerned with evolution. But the field of evolution is too wide and features that Abeetles examines will be described later. %Add in which section!!

The purpose of Abeetles is to be a prototype of scientific simulator. But not only for users from comunity of computer scientists, but also for those, who appreciate user friendly interface, even if it decreases performance of the system. 

\section{Term related to fields of concern of Abeetles}
%dopis, docisti!
Abeetles concerns with evolution. For writing about matters, that are concerned with it, several key terms should be included. Gene is a structural unit of inheritance in living organisms. A gene is, in essence, a segment of DNA that has a particular purpose, i.e., that codes for (contains the chemical information necessary for the creation of) a specific enzyme or other protein.\cite{GeneDic}
The distinction between phenotype and genotype is fundamental to the understanding of heredity and development of organisms. The genotype of an organism is the class to which that organism belongs as determined by the description of the actual physical material made up of DNA that was passed to the organism by its parents at the organism's conception. For sexually reproducing organisms that physical material consists of the DNA contributed to the fertilized egg by the sperm and egg of its two parents. For asexually reproducing organisms, for example bacteria, the inherited material is a direct copy of the DNA of its parent. The phenotype of an organism is the class to which that organism belongs as determined by the description of the physical and behavioral characteristics of the organism, for example its size and shape, its metabolic activities and its pattern of movement.
\cite{GenPhen}
In biology the genome of an organism is its whole hereditary information and is encoded in the DNA (or, for some viruses, RNA). This includes both the genes and the non-coding sequences of the DNA. The term was coined in 1920 by Hans Winkler, Professor of Botany at the University of Hamburg, Germany, as a portmanteau of the words gene and chromosome.\cite{GenomeDef}
%Furthemore should be explained: genome




%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Precise definition of targets of the system Abeetles}

%Objectives of this thesis have already been briefly mentioned in the first chapter. This chapter has for its objective precise description of these targets and explanation of context of individual bullets where necessary. 
%no.. these are targets of all thesis, but this chapter is about targets of simulator Abeetles.

%First, swift enumeration: Targets of this thesis are to design the simulator Abeetles, implement it and  introduce several experiments to be run on the simulator. %This is just shortcut of paragraph from the first chapter

%Abeetles: simulates evolution of agents in environment
The main target of Abeetles, as well as of other similar programs, is to simulate artificial life. But artificial life is rather a general idea, because to simulate all natural life as we know it on the Earth is with contemporary means impossible. Therefore every simulator chooses just a restricted subset. As stated above, Abeetles is concerned with evolution. But even this field is too wide. The next step is thus to specify what will be evolved and how will look like the world, where it will be evolved. In the following text the subject of evolution will be called agents and the world will be called environment. 

%Target of this kind of simulation: models not like anything natural, but despite it search for features, that would bring them closer to natural ones. 
Abeetles is one of simulators, where both environment and agents are very simplyfied abstractions that do not have any concrete model in the nature. Abeetles does not endeavor to simulate any existing organisms, but to explore features of natural processes like evolution or natural selection and others on organisms that are from their basis virtual. Interesting question, that arises from this kind of simulation of evolutional processes is, what should parameters of agents and environment, whose models have little in common with real life, look like, so as to cause, that the results and course of evolution would resemble patterns of natural organisms in their environment.  

%with specific features and in This chapter will focus on more detailed features of the system Abeetles, will explain their context and outline their targets.%The third - is it true?

Browsing in the Internet, one easily ascertains, that there are already many simulators of artificial life. Therefore it is necessary to define targets of Abeetles in the context of other artificial life simulators. Attributes of several of these "brother programs" will be described in details and afterwards features of environment and agents of Abeetles will be chosen through careful comparison. For this purpose will serve five existing simulators, that should exemplify: Avida, Bitozoa, Gene Pool, Mitozoos and Primordial life. Reasons for choice of each of them will be described later. Common attribute of all of them is that their binaries are freely or almost freely accesible. Abeetles is also free software and is intended to examine contributing set of features of natural life only within free software.
%Alternative:Accordingly Abeetles only strive to present an original set of features and attitudes among well-described parts of free and accesible implementations of artificial life.

%Already written at the beginning of the paragraph:
% Features of selected simulators will be described and on their base will be chosen the set of features of the nature Abeetles will focus on and other attributes of the simulator.

The simulators are classified by twenty-seven features. Features are sorted into   
six cathegories: Agents, Mating, Environment, User interface settings and implementation.

Cathegory Agents contains following features: Body(=description of body of agents), Motion(=how do agents move), Behavior(=what they do besides basic activities of life - motion, mating and eating), Features(=features that change during life), Life-span, Genes(=representation of genes), Phenotypic space(=space of all possible phenotypes of agents), Learning(=whether some ability to learn is included in model of agents)

Cathegory Mating comprises: Number of offspring in one reproduction, (Genetic algorithm(=algorithm used for creation of genes of a descendant), Choice of partner(=features that a partner must have to be chosen for mating), Conditions of mating(=what conditions must be satisfied before mating), Investment in offspring(=the amount of some resource, usually energy, that is given to a descendant by its parents at its birth)

Cathegory Environment includes: Description(=what does the environment look like), Features(=parameters of the environment, that are not represented by objects contained in the environment),Elements(=objects, that inhabit or are placed in the environment together with agents)

Cathegory User Interface Settings  encompasses attributes, by which can a user influence the artificial life run by the simulator. They are sorted in 3 cathegories, Mating, Life of individual and Environment.

Cathegory Statistics is alone.

Cathegory Implementation holds: Implementation(=features of implementation of the simulator), Distribution of computation(=whether and how it is possible to distribute the computation of the artificial life in the simulator to more computers), Speed settings(=whether and how it is possible to influence the speed of computation), Available sources(= whether source file of the simulator are available), Availability of binaries(=under what conditions are binaries available, e.g. on registration, free, ect.), Programming language(=programming language used and key libraries)

 %*****Opening paragraph about every simulator: 
 %Name and version that I examined, 
 %url address
 %author
 %cathegorisation into 3 systems
 
\section {Avida}

Avida is an often presented example of a program based simulator. It studies evolution of programs. For definition targets of Abeetles is important Avida's implementation - two programs, gui and console version - serve to scientific purposes, can run very fast, but can ofer graphical output as well. Examined release was Avida version 2.0b7 2003 for OS Windows. Avida is a joint project of the Digital Life Laboratory at the California Institute of Technology, \url{http://dllab.caltech.edu/}, (headed by Chris Adami) and the Digital Evolution Group, \url{http://devolab.cse.msu.edu/}, at Michigan State University. Web page of the project is \url{http://dllab.caltech.edu/avida/}.

\begin{tabular}{|p{150pt}|p{220pt}|}
\hline

&program based, evolution+celular automata, scientific simulator \\ \hline
textbf{Agents}&segments of code in a simple language called strings  \\ \hline
Body&one cell in the grid of the environment \\ \hline
Motion&no \\ \hline
Behavior&execution of the code \\ \hline
Features&content of stack and registers \\ \hline
Life-span&not restricted \\ \hline
Genes&string is itself a genome \\ \hline
Phenotypic space&any finite string constructed of instruction of the language. \\ \hline
Learning&no \\ \hline \hline
\textbf{Mating}& \\ \hline
Number of parents&1 \\ \hline
Number of offsprings&1 \\ \hline
Genetic algorith&Poisson-random mutation \\ \hline
Choice of partner&no \\ \hline
Conditions of mating&no \\ \hline
Investment in offspring&no \\ \hline \hline
\textbf{Environment}& \\ \hline
Description&a grid of cells in the form of torus \\ \hline
Features&resources and tasks, completion of a task gives a bonus to the string \\ \hline
Inhabitants&strings \\ \hline \hline
\textbf{UI settings}& \\ \hline
Mating&mutation rate \\ \hline
Life of individual&initial genomes of initial individuals \\ \hline
Environment&resources and tasks of the environment \\ \hline \hline
\textbf{Statistics}&instruction viewer, string details, etc. \\ \hline \hline
\textbf{Implementation}& \\ \hline
Distribution of computation&no \\ \hline
Speed settings&console version versus gui version  \\ \hline
Available sources&yes \\ \hline
Availability of binaries&free \\ \hline
Programming language&C++ \\ \hline

\end{tabular}

%*****Opening paragraph about every simulator: 
 %Name and version that I examined, 
 %url address
 %author
 %cathegorisation into 3 systems

\section {Bitozoa}
Bitozoa is a simulator created by M. Borkowski, that uses for representation of agents neural network in combination with a set of parameters. It evolves two types of agents, herbivores and carnivores. It was created for purpose of amusement and education. 
Used version is Bitozoa 2 - artificial life & artficial intelligence simulation, 2000. Web page of the project is \url{http://www.bpp.com.pl/bitozoa2/bitozoa2.html}.

\begin{tabular}{|p{150pt}|p{220pt}|}
\hline
&Bitozoa\\ \hline
&Neural networks, evolution, ?\\ \hline
textbf{Agents}&Bitozoa - herbivores(eat plants) and carnivores(eat herbivora)\\ \hline
Body&ball with 5 eyes, 2 flagella, various number of neurons, neural network with various topology. Eyes see all object in the world.\\ \hline
Motion&moves using flagella\\ \hline
Behavior&no\\ \hline
Features&energy level - for motion and work of neurons, burned with speed according to age \\ \hline
Life-span&(?)\\ \hline
Genes&(?)\\ \hline
Phenotypic space&angle of flagella connection to body, number of neurons, topology of neural network, sensitivity of eyes\\ \hline
Learning&no\\ \hline
\hline \textbf{Mating}&\\ \hline
Number of parents&2 or 1\\ \hline
Number of offsprings&1-n (2parents) nebo 1(1parent)\\ \hline
Genetic algorithm&random number of genes from both parents + mutation\\ \hline
Choice of partner&same species and high enough stamina\\ \hline
Conditions of mating&meeting of two bitozoa \\ \hline
Investment in offspring&both parents give the same fixed amount\\ \hline
\hline \textbf{Environment}&\\ \hline
Description&space in form of toroid\\ \hline
Features&number of spots, where food grows\\ \hline
Elements&food growing on special spots, herbivores and carnivores\\ \hline
\hline \textbf{UI settings}&\\ \hline
Mating&breed distance, breed level of stamina, inherited stamina\\ \hline
Life of individual&asexual reproduction allowed, costs of an action, flagella angle and efficiency, view of bitozoa, adding or killing of a bitozoa, stamina from food, strength of carnivore, initial values of: eye sensitivity, number of neurons and stamina\\ \hline
Environment&number of bitozoa and food spots at the beginning, time increment, viscosity, random numbers generator seed\\ \hline
\hline \textbf{Statistics}&graph of population, energy flow and energy flow averaged, simulation description\\ \hline
\hline \textbf{Implementation}&\\ \hline
Distribution of computation&no\\ \hline
Speed settings&animation on-off\\ \hline
Available sources&partial\\ \hline
Availability of binaries&permission from author\\ \hline
Programming language&C++ Win32API\\ \hline
&\\ \hline
do uvodu:&energy level - for motion and work of neurons, burned with speed according to age \\ \hline

\end{tabular}

\section {Gene Pool}
Gene Pool was created by Jeffrey Ventrella, current version is 5. It is a parameter based simulator, dealing with evolution of swimming creatures in a virtual Darwinian aquarium. It is a game with intriguing graphical interface, san serve as an entertaining learning tool, but author demonstrated with it also several interesting results concerning influence of atractiveness of partners on  results of evolution.\cite{GenePool1} \cite{GenePool2}
  
\begin{tabular}{|p{150pt}|p{220pt}|}
\hline
&Gene Pool\\ \hline
&Parameter based\\ \hline
textbf{Agents}&swimbot\\ \hline
Body&mouth, genitals and 2-10 parts for moving\\ \hline
Motion&swimming using parts, algorithm common for all swimbots\\ \hline
Behavior&nothing\\ \hline
Features&energy\\ \hline
Life-span&not restricted\\ \hline
Genes&(?)\\ \hline
Phenotypic space&length, width, system of attachment, phases and amplitudes of every part\\ \hline
Learning&no\\ \hline
\hline \textbf{Mating}&\\ \hline
Number of parents&2\\ \hline
Number of offsprings&1\\ \hline
Genetic algorithm&crossover algorithm with mutation\\ \hline
Choice of partner&"in round ""view horizon"" at one snapshot. Swimbot chooses at one snapshot in his ""view horizon"" one mate, that most satisfies his attractiveness criterion"\\ \hline
Conditions of mating&swimbots mate when at least one of them is pursuing the other and the distance between their genitals is less than the length of the genital vector\\ \hline
Investment in offspring&50\% of actual energy\\ \hline
\hline \textbf{Environment}&\\ \hline
Description&rectangular pool\\ \hline
Features&constant amount of energy in environment, cyling: food->swimbots->pool->food\\ \hline
Elements&food bits, swimbots\\ \hline
\hline \textbf{UI settings}&\\ \hline
Mating&attraction criterion\\ \hline
Life of individual&location in environment, creation a swimbot, hunger threshold, energy for offspring\\ \hline
Environment&food growth - birth delay, spread radius, energy from 1 piece\\ \hline
\hline \textbf{Statistics}&graph of number of swimbots vs. food, features of individual, the best one in attraction/mating/eating\\ \hline
\hline \textbf{Implementation}&\\ \hline
Distribution of computation&no\\ \hline
Speed settings&no\\ \hline
Available sources&no\\ \hline
Availability of binaries&yes\\ \hline
Programming language&(?)\\ \hline
&\\ \hline
do uvodu:&how good is a swimbot in reproducing, which means either being good at swimming or being attractive to other swimbots\\ \hline

\end{tabular}

\section {Mitozoos}
Author of project Mitozoos is Spanish software company Bestiario, \url{http://www.bestiario.org/}. Its purpose is educational, as author states "Mitozoos is an interactive artificial life model created with the objective that through experimentation and play participants will understand the relationship between genetic code and life." The web presentation of the project can be found at \url{http://bestiario.org/mitozoos/english/index.html}.

\begin{tabular}{|p{150pt}|p{220pt}|}
\hline
&Mitozoos \\ \hline
&Parameter based \\ \hline
textbf{Agents}&mitozoos \\ \hline
Body&spider-like, with 2 eyes, body in shape of tetrahydron and four bent legs \\ \hline
Motion&walking using legs, algorithm common for all mitozoos \\ \hline
Behavior&strategy of eating, frequency and duration of resting periods \\ \hline
Features&energy \\ \hline
Life-span&not restricted \\ \hline
Genes&ten genes, each gene has four bases (= colors) \\ \hline
Phenotypic space&length of parts of legs, procreation threshold, eating strategy \\ \hline
Learning&no \\ \hline
\hline \textbf{Mating}& \\ \hline
Number of parents&2 \\ \hline
Number of offsprings&1 \\ \hline
Genetic algorithm&crossover algorithm with mutation \\ \hline
Choice of partner&no \\ \hline
Conditions of mating&meeting of two mitozoos with energy higher than procreation threshold \\ \hline
Investment in offspring&a fixed amount of energy \\ \hline
\hline \textbf{Environment}& \\ \hline
Description&space in form of a circle \\ \hline
Features&rate of food growth \\ \hline
Elements&mitozoos, food bits \\ \hline
\hline \textbf{UI settings}& \\ \hline
Mating&no \\ \hline
Life of individual&special application, which avails user to create a mitozoos by setting its genes and add it to running life. \\ \hline
Environment&initial number of mitozoos and food pieces, food growing rate, mutation rate \\ \hline
\hline \textbf{Statistics}&graph of number of swimbots vs. food, features of individual, genetic landscape, log of events, on reproduction join crossjoin of genotypes is shown \\ \hline
\hline \textbf{Implementation}& \\ \hline
Distribution of computation&no, but coders can run on different computers. \\ \hline
Speed settings&no \\ \hline
Available sources&no \\ \hline
Availability of binaries&yes \\ \hline
Programming language&Flash, (?) \\ \hline
\end{tabular}

\section {Primordial Life}
For following evaluation Primordial Life 3.0 by Jason Spofford was tried. It can be downloaded at \url{http://www.io.com/~spofford/prim30.html}. Primordial life is a shareware parameter based artificial life screen saver writen for purpose to capture the principles of evolution in an interestion and visual way.

\begin{tabular}{|p{150pt}|p{220pt}|}
\hline
&Primordial life 3.21\\ \hline
&parameter based, evolution, game\\ \hline
textbf{Agents}&biots\\ \hline
Body&pattern of vectors of different colors, color of a part determines its special function.\\ \hline
Motion&moved by environment or by its motor (light blue parts)\\ \hline
Behavior&fighting with each other\\ \hline
Features&energy level\\ \hline
Life-span&restricted, constitution of body\\ \hline
Genes&64 numbers, describe lines color, orientation and length. Certain genes have special meaning detailing how many lines a biot has, its symmetry, whether it has mirrored or radial symmetry, how many children it should have and whether or not it should disperse its children after they are born. Complete description available.\\ \hline
Phenotypic space&number and symmetry of vectors, color, orientation and length of each vector, number and dispersion of offspring in a birth\\ \hline
Learning&learn from collision with other biots - change body accordingly\\ \hline
\hline \textbf{Mating}&\\ \hline
Number of parents&2 nebo 1\\ \hline
Number of offsprings&1-n\\ \hline
Genetic algorith&genetic crossover with mutation(2 parents), mutation (1 parent)\\ \hline
Choice of partner&no\\ \hline
Conditions of mating&collision of two biots, one of them with white parts\\ \hline
Investment in offspring&yes\\ \hline
\hline \textbf{Environment}&\\ \hline
Description&rectangular space\\ \hline
Features& mixing of biots continuely up and providing of light, absorbable by green parts of biots and used up as energy\\ \hline
Elements&biots\\ \hline
\hline \textbf{UI settings}&\\ \hline
Mating&mutation rate, sexual/asexual/ both reproduction\\ \hline
Life of individual&life span, speed, regeneration rate and cost, attacking of child, battle of sibblings, level of selfchange thanks to collision\\ \hline
Environment&solar intensity, friction, starting population, plague - its duration and lethality\\ \hline
\hline \textbf{Statistics}&ecosystem status (population, death and birth rate numbers), global status of all connected ecosystems(no. of connected ecosystems, population)\\ \hline
\hline \textbf{Implementation}&\\ \hline
Distribution of computation&automatically connecting ecosystems together over the Internet.\\ \hline
Speed settings&no\\ \hline
Available sources&no\\ \hline
Availability of binaries&register\\ \hline
&\\ \hline
&\\ \hline
do porovnaniu:&"  A Primordial Life Server provides lists of available ecosystems you can connect with.   Ecosystems still connect peer to peer& the server just provides addresses of other ecosystems willing to let you connect with them. Advanced setting available !!"\\ \hline
\end{tabular}

\section{Features of Abeetles}
"Brother programs" were compared according to 27 different cathegories. Abeetles does not strive to be contributing in all these cathegories. Instead of it, four were chosen as the key ones: motion, learning, choice of partner, distribution of computation and elements of environment. The following table compares Abeetles with other solutions specificaly in this cathegories.

\begin{tabular}{lcccccc}
 & Avida & PLife&  Bitozoa& Mit.& GPool & Abeetles\\ \hline
Motion: strategy&x&x&G&G&S&G\\
Learning&x&O&x&x&x&G/O\\
Choice of partner&x&x&x&x&O&G/O\\
Distribution of computation&x&Y&x&x&x&Y\\.
Elements of environmet: obstacles&x&x&x&x&x&O
\end{tabular}

Motion strategy is function of an organism, where input is the view and output is decision where to move.
O - optional (user can influence it)
S - static (fixed)
G - genetically evolved
x - not present



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Model of Abeetles}
%*******************1-2 pages about choice of my objectives according to previous chapter. I will use Porovnani.xls
Abeetles would like to contribute to the set of artificial life applications. Its predecesor is the program Broucci developed at the Faculty of Mathematic and Physics at Charles University in Prague by RNDr.Tomas Holan Phd. as a  sample program for teaching purposes. The source code of Broucci is available at \url{http://ksvi.mff.cuni.cz/\~holan/jinak/}.

Program Broucci implements interesting ideas - evolution of strategy of motion of simple beetles in a grid with obstacles or in case when beetles eat each other. Closer description of this application can be found in chapter Survey of existing artificial life applications. Abeetle overtook some of these ideas and strives to apply them in new ways.\\
On the whole, there are quite a few artificial life applications. This topic became famous in past decade and encouraged this field of knowledge to considerable development. Existing applications cover wide range of interesting features and access the problematics from different points of view. Some of these applications and articles on this topic work are accessible completely, with all documentation, other are poorly documented or without source code and some are even not accessible freely at all. For example the website of International Society of Artificial Life demands membership fees for access to presented information. Therefore the objective to allocate a set of features and attitudes for a new artifical life system that would be verifiably original is hardly reachable.

Accordingly Abeetles only strive to present an original set of features and attitudes among well-described parts of free and accesible implementations of artificial life.



\section{Chosen features}
    All the evaluated artificial life applications were compared in 27(TO CHECK) cathegories. For Abeetles four were chosen as the key ones. The following table compares Abeetles with other solutions specificaly in this cathegories.

\begin{tabular}{lcccccc}
 & Avida & PLife&  Bitozoa& Mit.& GPool & Abeetles\\
Motion strategy&x&x&G&G&S&G\\
mems learning&x&O&x&x&x&G/O\\
Expectations on partner&x&x&x&x&O&G/O\\
Distribution of computation&x&Y&x&x&x&Y\\.
Obstacles in evironment&x&x&x&x&x&O
 \end{tabular}

individual motion strategy = (input=view, output=moving strategy)
O-optional
S-static
G-genetically evolved
x-Not present
\indent

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Structure of the thesis in relation to chosen objectives}
%4.         Struktura práce ve vztahu k vytyèeným cílùm

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Analysis of the problem from different aspects}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Proposal of solution with discussion of alternatives}
%***********Návrh øešení s diskusí alternativ
%*******1. volba modelu a diskuse moznosti (vetsi nastavitelnost, co je geneticky vyvijeno a co je pevne dano, co je mozno v prubehu menit)
%*******2. volba implementacniho reseni - konzole, paralelizace, gui, davkove reseni


\section{Description of the model of a beetle}
Representation of an agent(beetle) in Abeetles from genetical point of view:
Chromosome of a beetle includes following genes:\\
\\
Brain [2][4][4][4] - Set of actions, each of them is chosen by the beetle in the proper situation. Situation is defined by the quaternion (hunger,what is in the left, the front, the right). Item "hunger" specifies whether or not is the beetle hungry. Values of what is in the three directions from the beetle are four: beetle, flower, wall or nothing. These four possibilities are precisely the four possible contents of cells of the environment.\\
Actions, for which the beetle can decide in this situations are chosen from the following set: make one step, turn right, turn left.\\
Originally there were five actions: these three and copulation and waiting. Decision for copulation has sense only in situations, where is another beetle in front of the beetle and when the beetle is not hungry. And, furthemore, it these situations the beetle should always or almost always copulate (add here more rigorous explanation)when he meets another beetle. As a result, it is compulsory action in case a beetle sees another beetle in front of him.\\
 The action waiting was omitted - it gets beetles into deadlock situation. They should never decide for waiting, when they are hungry, because there cannot grow any flower on the spot, where they are staying. And when they are not hungry, it might be useful only as waiting for somebody to come and make child with him. But even in this situation it is not clever - beetle cannot finish it, because it has no memory. So it waits until it gets hungry or situation around changes.\\
ExpectOnPartner - Age [2] = 2B how much older / younger can be the partner
ExpectOnPartner - Energy = 1B how much more than ExpectOnPartner - InvInChild
ExpectOnPartner - InvInChild = 1B how much more than InvInChild
ExpectOnPartner - LearningAbility [2]= how much less/more can have the parter
InvInChild = 1B
HungryThreshold = 1B
LearningAbility = 1B

\subsection{Description of the model of the environment}

\subsubsection{Motion strategy}
The predecessor of Abeetles, Broucci, evolved

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Detailed design of chosen solution}
% Detailni rozpracování modelu reseni - tj. datove struktury k implementaci

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Description of prototype implementation}
%Detailed description of both prototypes - GUI version and console application  - from technical point of view.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Resume of fullfiling of objectives and other known attitudes}
%******************Comparison of result with brother programs from 3rd chapter
%9.         Zhodnocení splnìní cílù a srovnání s jinými známými pøístupy

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Resume with brief summon of results and contribution}
%10.         Závìr se struèným shrnutím výsledkù a pøínosu


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{References to related web pages}


\url{http://myxo.css.msu.edu/cgi-bin/lenski/prefman.pl?group=al} - Evolution experiments with digital organisms



\chapter {Implementation}

\section {Paralelisation}

\subsection Idea of paralelisation in Abeetles:
Paralelisation should distribute the computation of the grid environment for abeetles on several computers. As the environment is a grid, it can be cut into rectangular parts, each of this executed and evolved on a different source computer. This parts can have versatile size according to performance of the computer. With this schema, the number of fields of one part, that are computed on one computer is up to quadratical to width(heigth) of the part and number of fields of one part, that need communication with other parts is linear (four times) to width(heigth) of the part.\\
Why paralelisation can make the computation quicker: Beetles are located in a grid, so the interactions between the grid cells are similar to those of a celular automata - each beetle interacts only with its nearest neighborhood.
Artificial systems such as Tom Ray's tierra \cite{Tierra} have opened the possibility of studying evolvable systems. The study of evolution in such an information-rich artificial environment requires ever larger and faster systems, and present systems are largely restricted by such limits.
Distributing overmultiple processors is not practical on a large scale, because of the non-local interaction between members of the tierran population. But the system Avida developed the idea of the environment towards distribution. Avida is based on an array of cells that interact only with their nearest neighbours and an update mechanism reminiscet of 2D cellular automata.\cite{Avida} \\
Nevertheless Avida system does not include distributed computation in its latest implementation, neither in older ones.
Abeetles should exploit ideas presented in the article about Avida, follow suggestions about local interaction and other improvements of Avida and implement them in a system, that will execute in a paralel way. But unlike Avida, Abeetles will be a parameter based system, not a program based.




\subsection Choice of tool for Abeetles:
There are two main tools used for paralelisation of computation. PVM(Paralel virtual Machine) and MPI(Message Passing Interface). Comparison of features has been published in an article \cite {PVM versus MPI}

    \subsubsection {Topology:}
    MPI provides higher level of the message-passing topology. In MPI a group of tasks can be arranged in a specific logical interconnection topology. PVM does not support such an abstraction, the programmer is to manually arrange tasks into groups with the desired communication organization.
    For Abeetles the concept of MPI would be more convenient. But the organization of squares as parts of a toroid should be simple enough to realize using PVM.

    \subsubsection {Fault Tolerance:}
    Fault tolerance is important for any long-running simulation.  PVM has supported a basic fault notification scheme. Using this scheme individual tasks can inform other tasks about faults and respond to faults without hanging or failing.

    \subsubsection {Resource Control:}
    PVM is inherently dynamic in nature. Computing resources can be added or deleted any time. In a long­running application in which the user occasionally wishes to attach a graphical front­end to view the computation's progress. Without virtual machine dynamics, the graphical workstation would have to be allocated during the entire computation. MPI lacks such dynamics and is, in fact, specifically designed to be static in nature to improve performance.
    Abeetles intends to be such an application - fast running, with rare user requests for graphical view of progress (possibly) or graphs and metrics of the course of the computation. Thus PVM should serve better for the case of Abeetles.

    \subsubsection {Process Control:}
    Ability to stop and start tasks in included both in MPI(v.2.0) and PVM.

    \subsubsection {Portability versus Interoperability:}
    MPI programs are only portable, which means that they can be ported from one architecture to another, compiled and executed without modification. PVM programs can be dealt in the same way, but moreover can also communicate with each other.

    \subsubsection{Context for Safe Communication:}
    Offered by both MPI and PVM (v.3.4.)

    \subsubsection{Speed:}
    MPI is expected to be faster within a large multiprocessor. Reason is MPI's simplicity.

\section {Genetical Algorithm}

\subsection {Choice of genetical algorithm for Abeetles}

Representation of an agent(beetle) in Abeetles from genetical point of view:
Chromosome of a beetle is the set of variables, representing the beetle:

Brain [2][4][4][4] = 128B

I will consider the individual brain decisions as noninteracting and availe placing of a crossover point anywhere after finished set of decisions

I want following variables to be logicaly noninteracting and thus it should be possible to place a crossover point anywhere among them. Hence I represent the expectations on a partner as relative divergence from other values.

ExpectOnPartner - Age [2] = 2B how much older / younger can be the partner
ExpectOnPartner - Energy = 1B how much more than ExpectOnPartner - InvInChild
ExpectOnPartner - InvInChild = 1B how much more than InvInChild
ExpectOnPartner - LearningAbility [2]= how much less/more can have the parter
InvInChild = 1B
HungryThreshold = 1B
LearningAbility = 1B

For implementation of crossover step the possible crossover points must be carefully chosen. The target is to find points, that availe separation of the genome into two noninteracting groups. Otherwise the crossover is not important for evolution and may be omitted.

Points can be placed between any pair of adjacent variables. Two created parts will be independent


Generally:
\begin{itemize}
  \item 1.selection
  \item 2.crossover
    \begin{itemize}
      \item One-point crossover: This algorithm seems to be useful for the case of Abeetles. I will implement it as the first.
      \item Two-point crossover: This algorithm seems to be useful for the case of Abeetles. I will implement it.
      \item "Cut and splice": This technique is not suitable for case of Abeetles, because created children have different length of chromosomes.
      \item Uniform Crossover and Half Uniform Crossover: These aproaches may also be useful.
      \item Crossover for Ordered Chromosomes
    \end {itemize}
  \item 3.mutation
\end {itemize}



Sources:
\begin{itemize}
    \item \url{http://en.wikipedia.org/wiki/Crossover_\%28genetic_algorithm\%29}
    \item \url{http://www.springerlink.com/content/b67qv704abd9mtgt/fulltext.pdf}
    \item \url{http://www.springerlink.com/content/b67qv704abd9mtgt/}
    \item \url{http://www.cs.cas.cz/~roman/ea.html}
    \item \url{http://wiki.matfyz.cz/wiki/AIL025}
    \item \url{http://en.wikipedia.org/wiki/Swarm_intelligence}
    \end{itemize}

\section {Statistics and overviews offered by Abeetles}
GenePool: graph of no. of swimbots vs. food, features of individual, the best one in attraction/mating/eating
Mitozoos: graph of no. of swimbots vs. food, features of individual, genetic landscape, log of events, on reproduction join crossjoin of genotypes is shown
Bitozoa: graph of populations, energy flow and energy flow averaged, Save simulation description, Collect information to .txt file
Primordial life: ecosystem status (population, time death and birth rate numbers), global status of all connected ecosystems(no. of connected ecosystems, population)
Avida: instruction viewer, string details, graphs of

GenePool: graph of no. of swimbots vs. food,
          features of individual,
          the best one in attraction/mating/eating
Mitozoos: genetic landscape,
          log of events,
          on reproduction join crossjoin of genotypes is shown
Bitozoa:  graph of populations,
          energy flow and energy flow averaged,
          Save simulation description,
          Collect information to .txt file
Primordial life: ecosystem status (population, time death and birth rate numbers),
          global status of all connected ecosystems(no. of connected ecosystems, population)
Avida: instruction viewer,
       string details,
       graphs of

\section {Memory of beetles}
Beetles without memory have no chance to vary their decision, if the situation around them is constant. And if the decision is anything else than a step, their repeat this until they get hungry, die or the situation changes. But in case their are already hungry and the situation is a for long time constant, they are determined to die. And in the range of whole environment, if there are few beetles in the environment, they become extinct.
The other problematic situation ocurrs, when two beetles meet and want to copulate with each other. They might make a child or more, but then they do not have enough energy to create another child and get into deadlock, until situation around them changes or one of them dies. (Ne.. tohle nevadi, protoze se zmeni situace z hladove na nehladovou. Postreh: obecne nema smysl, aby hladovi broucci mohli volit akci A_COPULATE Postreh2: A je nutne, aby broucek pri kazdem setkani s jinym brouckem provedl A_COPULATE) Solution: omit action A_COPULATE from list of possible actions and include it as "hardwired". So this action will be performed always when in front of a beetle is another beetle and no decision will be made in this case. In case the copulation did not create a child beetle, e.a. expectations were not met, the decision for the case "beetle in the front" will be made according to the Brain table.

Ways how to solve this problem:
1. Flowers could besides growing also die back. Such a solution would  cause the environment to vary in every step, even if beetle don't eat the flowers.
2. Beetles could get aditional memory, to determine a deadlock-like situation. This could be done in several different ways:
a) remember n past coords - a beetle which does not move is likely to die.
b) remember n past actions - a beetle which does some other action than the step more than 3 times (rotation) or more than once (unsuccessful copulation) does it wastefully.



\chapter{my Remarks}

\url{http://www.alife.org/ - International society for Artificial life}

\url{http://www.scs.carleton.ca/\~soma/biosec/readings/spafford-viruses.pdf} - Interesting essay about considerations whether viruses are forms of artificial life or not. Result: No.
Paper includes a worthwhile definition of life:
    \begin{itemize}
    \item Life is a pattern in space-time rather than a specific material object.
    \item Self-reproduction, in itself or in a related organism.
    \item Information storage of a self-representation.
    \item A metabolism that converts matter/energy.
    \item Functional interactions with the environment.
    \item Interdependence of parts.
    \item Stability under perturbations of the environment.
    \item The ability to evolve.
    \item Growth or expansion
    \end{itemize}

\chapter {Ideas}
\begin {itemize}
  \item Pocet rodicu > 2 ... jak to ovlivni evoluci?
\end{itemize}

\begin{thebibliography}{99}

\bibitem{Bedau}
Mark A. Bedau (November 2003). Artificial life: organization, adaptation and complexity from the bottom up. TRENDS in Cognitive Sciences. \url{http://www.reed.edu/\~mab/publications/papers/BedauTICS03.pdf}

\bibitem{Langton}
Chris G. Langton(1989) Artificial Life Addison-Wesley. \url{http://zooland.alife.org/}

\bibitem{AdamiBrown}
Chris Adami and Titus Brown (2000) "What is Artificial Life?" The Seventh International Conference on the Simulation and Synthesis of Living Systems, Reed College, Portland, Oregon, USA, 2000

\bibitem{BornhoffenLataud}
Stefan Bornhofen and Claude Lattaud (2006) Artificial Evolution, chapter Outlines of Artificial Life: A Brief History of Evolutionary Individual Based Models, pages 226-237, Springer, April 2006

\bibitem{GeneDic}
Gene. (n.d.). Columbia Electronic Encyclopedia. Retrieved July 14, 2007, from Reference.com website: \url{http://www.reference.com/browse/columbia/gene-ent}

\bibitem{GenPhen}
Lewontin, Richard, "The Genotype/Phenotype Distinction", The Stanford Encyclopedia of Philosophy (Spring 2007 Edition), Edward N. Zalta (ed.), \url{http://plato.stanford.edu/archives/spr2007/entries/genotype-phenotype/}.

\bibitem{GenomeDef}
Joshua Lederberg and Alexa T. McCray (2001). "'Ome Sweet 'Omics -- A Genealogical Treasury of Words". The Scientist 15 (7). \url{http://lhncbc.nlm.nih.gov/lhc/docs/published/2001/pub2001047.pdf}

\bibitem{Avida}
Avida: A Software Platform for Research in Computational Evolutionary Biology;
C. Ofria and C.O. Wilke, Artif. Life 10 (2004) 191-229.

\bibitem{Tierra}
Ray, T. S. 1991, "Evolution and optimization of digital organisms", in Billingsley K.R. et al (eds), Scientific Excellence in Supercomputing: The IBM 1990 Contest Prize Papers, Athens, GA, 30602: The Baldwin Press, The University of Georgia. Publication date: December 1991, pp. 489-531.

\bibitem{GenePool1}
Jeffrey Ventrella 1998, "Attractiveness vs. Efficiency: How Mate Preference Affects Locomotion in the Evolution of Artificial Swimming Organisms", Artificial Life VI 1998, MIT Press. \url{http://www.ventrella.com/Alife/Attractiveness/attractiveness_0.html}
 
\bibitem{GenePool2}
Jeffrey Ventrella, "Gene Pool: Exploring the Interaction Between Natural Selection and Sexual Selection", Chapter 4 in Artificial Life Models in Software, edited by Andrew Adamatzky and Maciej Komosinski, Springer, 2005

\bibitem{Mitozoos}
The book of the mitozoos. (\url{http://bestiario.org/mitozoos/english/pdf/mitozoos\_en.pdf}). Mitozoos World is a project created and developed by Bestiario [www.bestiario.org]. Website of the project (\url{http://bestiario.org/mitozoos/english/index.html}).

\bibitem{PVM versus MPI}
PVM and MPI: a Comparison of Features; G.A.Geist J.A.Kohl P.M. Papadopoulos

\end{thebibliography}

\end{document}
